---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
# Vectors
age = c(22, 21, 24, 19, 20, 23)
age
age[1]
yrs_math_ed = c(4, 5, 2, 5, 3, 5)
yrs_math_ed

# Data frames
df1 = data.frame(Age = age, Years = yrs_math_ed)
df1
df1[1,]
df1$Years

a = c("Mary","Martha","Kim","Kristen","Amy","Sam")
b = c("English","Math","Sociology","Math","Music","Dance")

df2 = data.frame(Name = a, Major = b)
df2

df3 = data.frame(Age = age, Years = yrs_math_ed, Name = a, Major = b)
df3

df4 = cbind(df1, df2)
class(df4$Name)
summary(df4)

d = c(19, 4, "John", "Math")

df5 = rbind(df4, d)

dfCopy = df4

df5[df5$Years > 4,]

df5[df5$Major == "Math",]

#Import data
Example1 = read.csv("~/Dropbox/Personal/SMU MSDS/03 - Doing Data Science/MSDS_6306_Doing-Data-Science/Unit 1/BusinessSales.csv", header = TRUE)
Example1$sales

Example2 = read.csv(file.choose(), header = TRUE)
Example2$sales

df = data.frame(x = c(1,2), y = c(3,4), z = c(5,6))
df[,2]

```

```{r}
# Plotting
Example1 = read.csv("~/Dropbox/Personal/SMU MSDS/03 - Doing Data Science/MSDS_6306_Doing-Data-Science/Unit 1/BusinessSales.csv", header = TRUE)
head(Example1)

plot(Example1$ad_tv, Example1$sales, pch = 15, xlab = "X", ylab = "Y", main = "Title")
abline(h = 55, col = "red", lwd = 5)

library(ggplot2)

plot(iris$Petal.Length, iris$Sepal.Length)

irisvir = iris[iris$Species=="virginica",]
plot(irisvir$Sepal.Length)

# Histograms
hist(mpg$cty, col = "blue")

# Boxplots
boxplot(cty~class, data = mpg)

# Dividing plots
par(mfrow = c(1,2))
hist(mpg$cty, col = "blue")
boxplot(cty~class, data = mpg)


dev.off()
hist(iris$Petal.Length)

barplot(df3$Years, names.arg = df3$Name)

summary(mpg$class)

mpg$classFact = as.factor(mpg$class)

barplot(summary(mpg$classFact))

pairs(~mpg+disp+drat+wt, data = mtcars)
```

```{r}
# Basic stats

sample1 = rnorm(1000, 0, 1)

hist(sample1)
mean(sample1)
sd(sample1)

population = rnorm(100000000, 0 ,1)

sample1 = sample(population, 100)
hist(sample1)
mean(sample1)
sd(sample1)

population = rnom(10000000,0,1)

xBarVec = c()
xBarGenerator = function(sampleSize = 30, number_of_samples = 100)
{
  for (i in 1:number_of_samples) {
    tempSample = sample(population, sampleSize)
    xBarVec = c(xBarVec, mean(tempSample))
  }
  return(xBarVec)
}

x=xBarGenerator(30, 1000)

length(x)

x

hist(x)

```

## You Try It
Now adapt the code above to generate a specified number of sample means of a specified size sample size (this has already been done) from a normally distributed population with a specified meanand standard deviation (currently it is forced to be from a standard normal distribuion N(0,1)).

Run it a few times and use it generate 10000 sample means, each from a sample of size 50, from a normal distribution with mean = 60 and sd = 10.  Provide a distribution (histogram) of the distrubtions as well. 

What is the shape of the distribuion?  What is the mean and sd?  What should they be given the Central Limit Theorem (CLT).  For a refresher of the CLT see the Stats Bridge Course!

```{r}

xBarGenerator = function(mean = 0, sd = 1, sampleSize = 30, number_of_samples = 100)
{
  for (i in 1:number_of_samples) {
    tempSample = rnorm(sampleSize, mean, sd)
    xBarVec = c(xBarVec, mean(tempSample))
  }
  return(xBarVec)
}

x=xBarGenerator(mean = 60, sd = 10, sampleSize = 50, number_of_samples =  10000)

length(x)

x
mean(x)
hist(x)



```

