---
title: "Unit 2 Visualization"
author: "Andrew"
date: "4/18/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

## UNIT 2 R Quick Start / Scatterplot, histogram and bar plot.R

### Remember Previous Plots from Unit 1
### Let's redo them using ggplot2!

Scatterplot
```{r}
#install.packages("ggplot2")
library(tidyverse)


# You Try It! (mpg)
# plot (scatter plot) city mile per gallon (y) versus highway miles per gallon (x).

mpg |> 
  ggplot(aes(x = hwy, y = cty)) +
  geom_point()+
  ggtitle("Some title")
  
mpg |>
  ggplot(aes(x = cty)) + 
  geom_histogram()

mpg |>
  ggplot(aes(x = class, y = cty)) +
  geom_boxplot()

mpg |>
  ggplot(aes(x = class)) + 
  geom_bar()

```


Histogram
```{r}
#Iris histogram

# You Try It!  investigate the distribuion of the city miles per gallon variable for compact cars.  

#You Try It!  No compare the distribution of cty for compact cars to that of SUVs.  

# You Try It Bonus!  In order to compare the distributions, we should put them on the same scale.
# Research how to change the scale!


```

Boxplot
```{r}
#Iris Boxplot

# You Try It!

#Find a box plot from the mpg dataset of the city mpg versus the class of car.

```



### Bar plot with mpg and class
```{r}
#With base barplot

# With ggplot /  geom_bar()

```

## Bar plot with toy Student and Math Experience data
```{r}
#With base barplot
age = c(22,21,24,19,20,23)
yrs_math_ed = c(4,5,2,5,3,5)
names = c("Mary","Martha","Kim","Kristen","Amy","Sam")
subject = c("English","Math","Sociology","Math","Music","Dance")

df3 = data.frame(Age = age, Years = yrs_math_ed, Name = names, Subject = subject)

# With ggplot /  geom_bar()   Where does Years go? We will come back to this!

df3 |>
  ggplot(aex(x))

# You Try It!
# Iris Data:  Create a Barplot of how many of each Species are in the study.



# You Try it Bonus: Create a dataframe of how you rank (scale of 1(least) to 10 (best)) traveling by car, train, plane, boat (ship), and bike.


```



## AES
### AES: X and Y
```{r}
mpg |>
  ggplot(aes(x = class, y = ..prop.., group = 1)) +
  geom_bar()

mpg |>
  ggplot(aes(x = hwy, y = cty, color = drv)) +
  geom_point()

mpg |>
  ggplot(aes(x = hwy, y = cty, color = drv)) +
  geom_point(color = "blue")

mpg[mpg$class == "compact",] |>
  ggplot(aes(x = cty, fill = drv)) +
  geom_histogram()

summary(mpg)
  
mpg |>
  dplyr::filter(class == "compact") |>
  ggplot(aes(x = cty, fill = drv)) +
  geom_histogram()

mpg |>
  ggplot(aes(x = cty, fill = drv)) +
  geom_histogram(color = "black")

mpg |>
  ggplot(aes(x = class, y = cty, fill = drv)) +
  geom_boxplot(color = "black")

mpg |>
  ggplot() +
  geom_point(aes(x = cty, y = hwy, color = drv))

mpg |>
  ggplot(aes(x = displ, y = hwy, color = drv)) + 
  geom_point()

```

### AES: color
```{r}

# color = "blue" is not a mapping of a variable

#note that color is not in an ”aes”
# function we are setting them manually.

# note now we are ”mapping” fill to a variable (drv).

#note that fill and color are not in an ”aes” function we are setting them manually.
# We are not “mapping” variables to colors… rather just setting the color.  

# note now we are ”mapping” fill to a variable (drv).

#note that fill and color are not in an ”aes” function we are setting them manually.
# We are not “mapping” variables to colors… rather just setting the color.  


```


### AES: size
```{r}

```


### AES: size
```{r}


```


### AES: size
```{r}


```



### AES: alpha
```{r}


```

### AES: shape
```{r}


```


## GEOMS

### geom_smooth
```{r}


```

### geom_smooth you try it
```{r}


#Global aes() ... one aes that applies to all layers

#Individual aes() …. an aesthetic for each layer

#Individual aes() …. an aesthetic for each layer

mpg |>
  ggplot(aes(x = hwy, y = cty)) + 
  geom_point(aes(color = class)) +
  geom_smooth(aes(linetype = drv))

mpg |>
  ggplot(aes(x = hwy, y = cty, color = class)) + 
  geom_point() +
  geom_smooth() +
  facet_wrap(~class)

mpg |>
  ggplot(aes(x = hwy)) +
  geom_histogram()

```

### labels
```{r}

mpg |>
  ggplot(aes(x = class, y = cty)) +
  geom_boxplot() +
  ggtitle("XX") +
  xlab("X") + 
  ylab("Y")

```


### facets
facet wrap
```{r}

mpg |>
  ggplot(aes(x = hwy, y = cty, color = class)) +
  geom_point() + 
  geom_smooth() +
  facet_wrap(~class)


```

facet grid
```{r}
mpg |>
  ggplot(aes(x = hwy, y = cty, color = class)) +
  geom_point() + 
  geom_smooth() +
  facet_grid(drv~class)


```

## Transformations
```{r}

#stat = identity adds the values of the y value per class (x level)

mpg |>
  ggplot(aes(x = class)) +
  geom_bar(stat = "count")

mpg |>
  ggplot(aes(x = class, y = hwy)) +
  geom_bar(stat = "identity")

mpg |>
  ggplot(aes(x = class, y = hwy)) +
  geom_point()

```

## Positions
### stack
```{r}

mpg |>
  ggplot(aes(x = class, fill = drv)) +
  geom_bar(position = "fill")

```

### dodge
```{r}

mpg |>
  ggplot(aes(x = class, fill = drv)) +
  geom_bar(position = "dodge")

```

### fill
```{r}

```

### jitter
```{r}

mpg |>
  ggplot(aes(x = hwy, y = cty)) +
  geom_point(position = "jitter")

mpg |>
  ggplot(aes(x = hwy, y = cty)) +
  geom_point()

```

## Other Coordinate Systems
### coord_flip
```{r}

mpg |>
  ggplot(aes(x = class, fill = class)) +
  geom_bar() +
  coord_flip()

```

### coord_polar
```{r}

mpg |>
  ggplot(aes(x = class, fill = class)) +
  geom_bar() +
  coord_polar()

```

### coord_quickmap and map_data
```{r}
install.packages("maps")
library(maps)

usa = map_data("usa")

usa

p <- ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group), fill = "blue", color = "black") +
  coord_quickmap()
p

#Dallas Coords
Dallas <- tibble(
  long = c(-96.7970),
  lat = c(32.7767),
  names = c("Dallas")) 

p + geom_point(data = Dallas, aes(x = long, y = lat), shape = 21, color = "black", fill = "yellow", size = 5) +
  geom_text(data = Dallas, aes(x = long, y = lat, label = names), hjust = 0, nudge_x = 2, color = "white")


#Italy Map
italy = map_data("italy")
ggplot(italy, aes(long,lat, group = group)) + 
geom_polygon(fill = "white", color = "black") +
coord_quickmap()

?map_data
map_data(maps::)
```

## themes
```{r}
install.packages("ggthemes")
library(ggthemes)

#excel
mpg |>
  ggplot(mapping = aes(x = hwy, y = cty)) + 
  geom_point(position = "jitter") + 
  ggtitle("City MPG v. Highway MPG") + 
  theme_excel()

mpg %>% 
  ggplot(mapping = aes(x = class, fill = class)) +  
  geom_bar() + 
  theme_excel()

#Economist
mpg |>
  ggplot(mapping = aes(x = hwy, y = cty)) + 
  geom_point(position = "jitter") + 
  ggtitle("City MPG v. Highway MPG") + 
  theme_economist()

mpg %>% 
  ggplot(mapping = aes(x = class, fill = class)) +  
  geom_bar() + 
  theme_economist()


#538
mpg |>
  ggplot(mapping = aes(x = hwy, y = cty)) + 
  geom_point(position = "jitter") + 
  ggtitle("City MPG v. Highway MPG") + 
  theme_fivethirtyeight()

mpg %>% 
  ggplot(mapping = aes(x = class, fill = class)) +  
  geom_bar() + 
  theme_fivethirtyeight()

mpg |>
  ggplot(mapping = aes(x = hwy, y = cty)) + 
  geom_point(position = "jitter") + 
  ggtitle("City MPG v. Highway MPG") + 
  theme_bw()
mpg %>% 
  ggplot(mapping = aes(x = class, fill = class)) +  
  geom_bar() + 
  theme_bw()

```

## Plotly!!!
```{r}

install.packages("plotly")

#MPG
library(ggthemes)
library(plotly)

p = mpg %>% 
ggplot(mapping = aes(x = hwy, y = cty)) + 
geom_point(position = "jitter") + ggtitle("City MPG v. Highway MPG") + theme_excel()

p
 
ggplotly(p)


p = mpg[mpg$class == "compact",] %>% 
ggplot(aes(x = cty, fill = drv)) + 
geom_histogram(color = "blue") 

p

ggplotly(p)

mpg |>
  ggplot(mapping = aes(x = hwy, y = cty)) + 
  geom_point(position = "jitter") +
  ggtitle("City MPG v. Highway MPG") + 
  theme_excel()
```

Student Data
```{r}
#generating the data
age = c(22,21,24,19,20,23)
yrs_math_ed = c(4,5,2,5,3,5)
names = c("Mary","Martha","Kim","Kristen","Amy","Sam")
subject = c("English","Math","Sociology","Math","Music","Dance")
df3 = data.frame(Age = age, Years = yrs_math_ed, Name = names, Subject = subject)

#plotting code
p = df3 %>% 
ggplot(aes(x = Name, y = Years)) + 
geom_bar(stat = "identity") + 
ggtitle("Distribution of Class")

#enabling interactive piece
ggplotly(p)
```

## 3D Plots With Plotly
```{r}
#mpg
p <- plot_ly(mpg, x = ~cty, y = ~hwy, z = ~displ, color = ~drv) %>%
   add_markers() %>%
   layout(scene = list(xaxis = list(title = 'City MPG'),
                      yaxis = list(title = 'Displacement'),
                      zaxis = list(title = 'Highway MPG')))
p

# You Try... Create a 3D plot with plotly with x = sepal width, y = sepal.length, z = petal.width




p <- plot_ly(iris, x = ~Sepal.Width, y = ~Sepal.Length, z = ~Petal.Width, color = ~Species) %>%
   add_markers() %>%
   layout(scene = list(xaxis = list(title = 'Sepal Width'),
                      yaxis = list(title = 'Sepal.Length'),
                      zaxis = list(title = 'Petal.Width')))
p
```



